---
title: "p8105_hw1_kjw2175"
author: "Kennedy Wade"
date: "2025-09-19"
output: html_document
---
## Problem 0.1 / 0.2

Use R Markdown to write reproducible reports, GitHub for version control, and R Projects to organize your work.Create a public GitHub repo + local R Project,create a single .Rmd file named p8105_hw1_YOURUNI.Rmd that renders to github_document, submit a link to your repo via Courseworks.


## Problem 1

This “problem” focuses on correct styling for your solutions to Problems 1 and 2. 

```{r}
 # load packages
library(moderndive)
library(dplyr)
library(ggplot2)

# load the dataset
data("early_january_weather")

#mean temperature
mean_temp = mean(pull(early_january_weather, temp))

# scatterplot: temp vs time_hour, colored by humidity
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(
    title = "Temperature vs Time ",
    x = "Hour",
    y = "Temperature (°C)",
    color = "Humidity (%)"
  ) 

# export plot
ggsave("weather_scatterplot.png")


```

The `early_january_weather` dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.The variables names (columns) included are origin, year, month, day, hour, temp, dewp, humid, wind_direction, wind_speed, wind_gust,precipitation, pressure, visb, and time_hour. The mean temperature during this period was `r round(mean_temp, 2)` °C.The weather scatter plot shows that temperatures are generally lower in the early morning hours, rise throughout the day, and peak in the afternoon. Humidity tends to be higher when temperatures are higher and decreases slightly as temperatures decrease. The points are fairly evenly spread, and the color gradient effectively highlights changes in humidity throughout the day.


## Problem 2

This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

Let's create a dataframe.

```{r}
library(tidyverse)  

hw1_df <- tibble(
  vec_numeric = rnorm(10),                     
  vec_logical = vec_numeric > 0,                  
  vec_char = letters[1:10],                       
  vec_factor = factor(rep(c("1", "2", "3"), length.out = 10)) 
)

#Trying to take the mean of each variable using pull
mean(pull(hw1_df, vec_numeric))
mean(pull(hw1_df, vec_logical))
```

Now, convert variables to numeric using `as.numeric`

```{r}
# Convert variables to numeric
as_numeric_logical <- as.numeric(pull(hw1_df, vec_logical)) 
as_numeric_char <- as.numeric(pull(hw1_df, vec_char))   
as_numeric_factor <- as.numeric(pull(hw1_df, vec_factor)) 
```

When I first tried to take the mean of each variable in my dataframe, found that it worked for numeric and logical variables but did not work for character or factor variables. The character code failed because text values cannot be averaged and factors failed because factors are categorical and mean () does not work on them directly. After applying the `as.numeric()`code, logical values turned into 0s and 1s, character values became NA, and factors turned into numbers that represent their levels. This is why only numeric and logical variables can be averaged directly.









